# Univariate data {#univ}

In general, generating univariate data is relatively straightforward. `R` is full of functions that allow to generate data given a prespecified distribution, of which only the parameters have to be determined. As one of the simplest examples, consider the variable $X$, which follows a normal distribution with mean $\mu$ and variance $\sigma^2$
$$
X \sim \mathcal{N}(\mu, \sigma^2),
$$
such that
$$
f(X) = \frac{1}{\sqrt{2\pi\sigma^2}} \exp \Bigg{\{} - {\frac{(X-\mu)^2}{2\sigma^2}} \Bigg{\}}.
$$

In `R`, one can easily sample from this distribution, using the `rnorm` function that is in-built, with parameters `n`, `mean` and `sd`. In practice, if we want to generate $n = 1000$ values from a normal distribution with mean $\mu = 0$ and standard deviation $\sqrt{\sigma^2} = \sigma = 1$. Hence, $X$ is said to follow a standard normal distribution.

```{r}
set.seed(42)
n    <- 1000
mean <- 0
sd   <- 1
X <- rnorm(n, mean, sd)
```

It is simply verified that $X$ indeed seems to follow a univariate normal distribution by asking for some summary statistics.

```{r}
summary(X)
```

Additionally, we can plot the histogram of the variable $X$ to assess its shape.

```{r, message = F}
library(tidyverse)
library(ggthemes)
ggplot(mapping = aes(x = X)) +
  geom_histogram(aes(y = ..density..)) +
  stat_function(fun = dnorm, args = list(mean = mean, sd = sd)) +
  theme_minimal()
  
```





